Word-Referent Identification Under Multimodal Uncertainty

### 2018-04-13 Writeup/ms copy.Rmd
changes on ms

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

@@

```{r load_packages, include = FALSE}
library(papaja)
library(readr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(knitr)
library(boot)
library(dplyr)
library(langcog)
theme_set(theme_bw())
library(broom)
#library("knitcitations")
#cleanbib()
#options("citation_format" = "pandoc")
```

@@

How does the mind deal with uncertainty to identify the speaker's intended word? To address this question, many researchers in recent years have adopted the rational approach to cognition [@anderson90], which corresponds to Marr's computational level of analysis. This approach has had a tremendous impact not only on speech related research [@Norris08; @clayard08; @feldman2009; @kleinschmidt2015], but also on many other disciplines in the cognitive sciences [for reviews, see @chater06; @Knill04; @tenenbaum11]. Within this framework, researchers typically consider an "ideal observer" model which provides a characterization of the task/goal and shows what the optimal performance should be under this characterization. When there is uncertainty in the input, the ideal observer performs an optimal probabilistic inference. For example, in order to recognize an ambiguous linguistic input, the model uses all available probabilistic knowledge in order to maximize the accuracy of this recognition. Thus, when the task is well specified, the ideal observer model can be seen as a theoretical upper limit on performance. It is often used, not so much as a realistic model of human performance, as much as a baseline against which human performance can be compared [@Geisler2003]. When there is a deviation from the ideal, this can reveal extra constraints on human cognition, such as limitations on working memory or attentional resources.

In this line of research, @clayard08 simulated auditory uncertainty by manipulating the probability distribution of a cue (VOT) that differentiated similar words (e.g., "beach" and "peach"). They found that humans were sensitive to these probabilistic cues and their judgments closely reflected the optimal predictions. In another work, @feldman2009 studied the perceptual magnet effect, which is related to how categories influence the way the stimuli are perceived. They showed that this effect can be explained as the consequence of optimally solving the problem of perception under uncertainty. Both studies explored optimal performance under uncertainty from one modality, i.e., the auditory modality. There is however extensive evidence that information from the visual modality, in particular the speaker's facial features,  influences the perception of auditory information [see @Campbell2008 for a review]. @bejjanki2011 offered a mathematical characterization of how cues from speech and lip movements can be optimally combined and showed that human performance during audio-visual phonemic labeling was consistent (at least at the qualitative level) with the behavior of the ideal observer.

The speaker's facial features are not the only visual information that can be used to disambiguate the auditory input, however. For instance, in a referential context, information about the identity of the semantic referent can be integrated with linguistic information to resolve lexical and syntactic ambiguities in speech [e.g., @Eberhard1995; @Tanenhaus1995; @spivey2002]. To our knowledge, however, no study offered an ideal observer analysis of word identification in such context, that is, when the mind has to combine cues from the sound and the referent. Imagine, for example, that someone is uncertain whether they heard "pea" or "bee", does this uncertainty make them rely more on the referent (e.g., the object being pointed at)? Vice versa, if they are not sure if they saw a bee or a fly, does it make them rely more on the sound? More importantly, when input in both modalities is uncertain to varying degrees, do they weight each modality according to its relative reliability (which is the optimal strategy), or do they over-rely on a particular modality (which is a sub-optimal strategy)?

The visual information provided by the referent is arguably different from the visual information provided by the speaker's facial features (such as lip movements). There seems to be at least two fundamental differences between these two cases, and both can influence the cue combination strategy:

First, in the case of audio-visual speech, modalities are understood to differ only in terms of their reliability. In a referential context, however, modalities also differ in terms of their roles in the referential process: the auditory input represents the symbol (i.e., the word) whereas the visual input represents the referent. It has been suggested that because of its referential property, speech is a privileged signal for humans, starting in infancy [see @vouloumanos2014 for a review].\footnote{There is a debate over whether speech is privileged for infant and adults for the same reasons. Whereas Waxman and colleagues suggest that speech is privileged for both infants and adults because of its ability to refer, Sloutsky and colleagues suggest that speech might not have a referential status from the start. Rather, according to Sloutsky and colleagues, speech seems to be prefered by infants only because of a low level auditory "overshadowing".} Thus, in a referential context, it is possible that we do not treat the auditory and visual modalities as equivalent sources of information. Instead, there could be a bias for the auditory modality beyond what is expected from informational reliability alone.

<!--An example of such integration is when we determine the size of an object using the visual and haptic modalities [@ernst02], or when we determine the spacial location of a stimulus using the visual and the auditory modalities [@alais04] -->
Second, in the case of audio-visual speech, the associations are expected to be perceptually correlated. The expectation for this correlation is such that when there is a mismatch between auditory and visual input, people still integrate them into a unified (but illusory) percept [@mcgurk1976]. In the case of referential language, however, the multimodal association is by nature *arbitrary* [@saussure1916; @greenberg1957].  For instance, there is no logical/perceptual connection between the sound "bee" and the corresponding insect. Moreover, variation in the way the sound "bee" is pronounced is generally not expected to correlate perceptually with variation in the shape (or any other visual property) in the category of bees. Thus, one could imagine that cue combination in the case of arbitrary audio-visual associations (word-referent) is less automatic, more effortful, and therefore less conducive to optimal integration than it is in the case of perceptually correlated associations (as in audio-visual speech perception).

In the current study, we investigate how people combine cues from the auditory and the visual modality to recognize words in a referential context. In particular, we study how this combination is performed under various degrees of uncertainty in both the auditory and the visual modality.  We perform a rational analysis of the task. First we propose an ideal observer model that performs the combination in an optimal fashion. Second we compare the predictions of the optimal model to human responses. Humans can deviate from the ideal for several reasons. For instance, as mentioned above, a sub-optimality can be induced by the suggested privileged status of speech or by the arbitrariness of the referential association.  In order to study possible patterns of sub-optimality, we compare the optimal model (which provides a normative benchmark) to a descriptive model which is identical to the optimal model, except that the parameters in the former are fitted to human responses. <!--We do not perform the ideal observer analysis at the individual level, but at the population level. There might be some individual variation, but this study addresses the overall performance in the population as a whole.-->

@@

# Models and predictions

In this section we describe the multimodal cue combination model. First, we briefly introduce the experimental paradigm, and second we explain how behavior in this paradigm can be characterized optimally with an ideal observer model.

@@

We introduce a new task that tests word recognition in a referential context.  We use two visual categories (cat and dog) and two auditory categories (/b/ and /d/ embedded in the minimal pair /aba/-/ada/). For each participant, an arbitrary pairing is set between the auditory and the visual categories, leading to two audio-visual word categories (e.g., dog-/aba/, cat-/ada/). In each trial, participants are presented with an audio-visual target (the prototype of the target category), immediately followed by an audio-visual test stimulus (Figure\ \@ref(fig:task)). The test stimulus may differ from the target in both the auditory and the visual components.  After these two presentations, participants press "same" or "different".

@@

This task is similar to the task introduced by @sloutsky2003, and used in subsequent studies in both children and adults [see @robinson2010 for a review]. However, in this original task, participants are asked whether or not the two audio-visual presentations are *identical*. Such task would allow us to probe audio-visual encoding, but not necessarily language understanding. The latter requires---in addition to the perceptual encoding---a categorical judgement, i.e, determining whether or not two similar tokens are members of the same phonological/semantic category. Thus, our task is category-based: Participants are asked to press "same" if they think the second item (the test) belonged to the same category as the first (target) (e.g.,  dog-/aba/), even if there is a slight difference in the sound, in the referent, or in both. They are instructed to press "different" only if they think that the second stimulus was an instance of the other word category (cat-/ada/). The task also includes trials where pictures are hidden (audio-only) or where sounds are muted (visual-only). These unimodal trials provide us with the participants' evaluation of the probabilistic information present in the auditory and visual categories. This unimodal distributions are used as inputs to the optimal combination model, described below.

<!--We assume, for the sake of simplicity, that category tokens vary along a single dimension. Phonetic tokens are sampled from an acoustic continuum linking /aba/ to /aba/, and obtained by continuously varying the value of the second formant [@vroomen2004]. Visual tokens are sampled from a morph continuum linking the picture of a dog and the picture of a cat [@freedman2001]. -->

@@

<!--First we describe how probabilistic representations in each modality are derived, and second we explain how they are combined to yield optimal predictions. -->

<!--
### Unimodal condition

In each modality, we have two categories: /ada/ ($A=1$) and /aba/ ($A=2$) in the auditory dimension, and *cat* ($V=1$) and *dog* ($V=2$) in the visual dimension.

We assume a token's probability of membership in each category to be normally distributed:

$$ p(a | A) \sim  N(\mu_A, \sigma^2_A) $$
$$ p(v | V) \sim  N(\mu_V, \sigma^2_V) $$
The parameters of these distributions are fitted to the participants' judgement in the unimodal case, i.e., when they only hear the sounds, and when they only see the objects. For an ideal recognizer, the probability of choosing category 2 (that is, to answer "different") when presented with, say, an audio instance $a$, is the posterior probability of this category $p(A_2|a)$. If we assume that both categories have equal variances, the posterior probability reduces to:

$$p(A_2 | a)=\frac{1}{1+(1+\epsilon_A)\exp(\beta_{a0}+\beta_aa)}$$

with $\beta_a=\frac{\mu_{A_1}-\mu_{A_2}}{\sigma^2_{A}}$, $\beta_{a0}=\frac{\mu^2_{A_2}-\mu^2_{A_1}}{2\sigma^2_{A}}$ and $1+\epsilon_A=\frac{p(A_1)}{p(A_2)}$ is the proportion of the prior probabilities. If the identity of the auditory categories is randomized, and if $A_1$ is the target, then $\epsilon$ measures a response bias to "same" if $\epsilon > 0$, and a response bias to "different" if $\epsilon < 0$.

We modeled the visual representation in exactly the same way.

The ideal observer model combines probabilistic information from the auditory and visual modalities, and makes a categorical decision. How can this task be characterized in a formal way, and what would be the optimal behavior? If the stimuli were continuous (e.g., size, location), cues derived from each modality would be characterized with their sensory noise variance only, and (assuming conditional independence) their combined information would be characterized with a simple linear combination rule where the combined estimate is an average of the individual estimates, each weighted according to its relative reliability [@ernst02; @alais04]. However, in our case decision is made over categorical variables (phonemes and concepts), the optimal model should take into account, not only the noise variability around an individual perceptual estimate, but also its categorical variability (i.e., the probability with which the estimate belongs to a given category).

One way to account for the categorical variability is to model the decision step which maps sensory input onto task-dependent categories [@Bankieris17; @bejjanki2011]. We represent the decision step as follows (see Figure\ \@ref(fig:model) for illustration). If we assume that both the visual and the auditory cues vary along a single dimension, an audio-visual signal $(a,v)$ can be represented as a point in this two-dimensional space, and a word category can be characterized with a bivariate probability distribution over audio-visual tokens.\footnote{For the sake of clarity, we assume here that the sensory noise around this point is small/negligible compared to the categorical variability. But see below how the noise can be integrated when its effect is relatively large} Each category is peaked around the most typical audio-visual token, and the probability of membership in the category (the likelihood probability) decreases gradually as the auditory token deviates from the typical sound instance (depending on the variance of the auditory category) and/or as the visual token deviates from the typical visual instance (depending on the variance of the visual category). <!--To illustrate how categorical decision is made within this formal framework, suppose we receive an ambiguous image which can be either of a cat or of a dog, associated with a slightly noisy auditory input that sounds more like "cat". If the judgment is based on the picture only, this would lead to a random choice between cat and dog. However, if the judgement is based on both the picture and the sound (which is less ambiguous), the combined information should lead to a preference for the word cat  we illustrate the cateorical decision making. Audio-visual points along this line lead to random responses, meaning that both the auditory and the visual components are maximally ambiguous. Audio-visual points that lay on either side of the decision threshold lead to decisions in favor of one word or the other. This means that at least one component of this audio-visual token is not totally ambiguous, which leads to a diambiguated combined information. -->

A crucial feature of the ideal observer model is that it makes use of the entire probability distribution of a given category in order to make optimal decisions [e.g., @clayard08]. This means that the probability of picking a word (the posterior probability) should mirror closely the probability of the combined information (the likelihood probability). For instance, the model becomes less ambiguous and more certain as the tokens get closer to the typical instance of the category. In what follows, we formalize this sense of optimality using a Bayesian framework.
<!--characterized with a likelihood function, e.g., $P(S_V | S)$ and $P(S_A | S)$ for a visual and an auditory cue, respectively. If these cues are conditionally independent, their combined information can be written as $P(S_V, S_A | S)=P(S_A | S)P(S_V | S)$. The maximum likelihood estimate of this combined information $\hat{S}$ is an average of the individual maximum liklihoods $\hat{S}=w_V\hat{S_V}+w_A\hat{S_A}$, each wighted according to its relative reliability:
$w_V=\frac{1/\sigma_V^2}{1/\sigma_V^2 + 1/\sigma_A^2}$, and $w_A=\frac{1/\sigma_A^2}{1/\sigma_V^2 + 1/\sigma_A^2}$.

If, however, the target is a categorical dimension, the optimal model should take into account, not only the noise variability arround an individual perceptual estimate, but also its categorical variability (i.e., the probability with which this estimate belongs to a given category). For example, upon hearing an audio-visual signal (the sound "cat" associated with the image of a cat), the ideal obsever has both to identify the sensory input and to map it onto the task-dependent categories.

The optimal model that takes into account categorical variability is somewhat different, since -->

<!-- more adequate to the case of redundant information, where, for instance, cues are in the same units, the same coordinates and about the same aspect of the environmental property. It is less adequate to describe the case of arbitrary associations where cues are in different units, coordinate systems, or about complementary aspects of the same environmental property [see @ernst04 for further detail]. -->

An auditory category $A$ and a visual category $V$ are defined as distributions in the auditory and the visual dimensions, respectively. A word category $W$ is defined as a bivariate distribution in the audio-visual. More specifically, we define it as the joint distribution of the auditory and the visual categories.  For simplicity, we assume both auditory and visual categories to be normally distributed. That is, if $w=(a,v)$ is an audio-visual token, then:
$$ p(a | A) \sim  N(\mu_A, \sigma^2_A) $$
where $\mu_A$ and $\sigma^2_A$ are respectively the mean and the variance of the auditory category,
$$ p(v | V) \sim  N(\mu_V, \sigma^2_V) $$
where $\mu_V$ and $\sigma^2_V$ are  the mean and the variance of the visual category, and

$$ p(w | W) \sim  N(M_W, \Sigma_W) $$
where $M_W=(\mu_A, \mu_V)$ and $\Sigma_W$ are mean and the covariance matrix of the word category. We assume that auditory and visual variables are independent (i.e., uncorrelated), so the covariance matrix is simply:
 \[
   \Sigma_W=
  \left[ {\begin{array}{cc}
   \sigma^2_A & 0 \\
   0 & \sigma^2_V \\
  \end{array} } \right]
\]

```{r model, fig.cap = "Illustration of the model using simulated data. A word category is defined as the joint bivariate distribution of an auditory category (horizontal, bottom panel) and a visual semantic category (vertical, left panel). Upon the presentation of a word token $w$, participants guess whether it is sampled from the word type $W_1$ or from the word type $W_2$. Decision threshold is where the guessing probability is 0.5.", fig.align = "center", out.width = "400px"}
knitr::include_graphics("pictures/model.png", dpi = 108)
```

@@

Now we turn to the crucial question of modeling how optimal decision should proceed based on the probabilistic information in the auditory and the visual modalities. In our task, we have two word categories: dog-/aba/ ($W_1$) and cat-/ada/ ($W_2$).\footnote{This mapping is randomized in the experiments.} Participants can be understood as choosing one of these two word categories (Figure\ \@ref(fig:model)). For an ideal observer, the probability of choosing category 2 when presented with an audio-visual instance $w=(a,v)$ is the posterior probability of this category:
$$
p(W_2 | w)=\frac{p(w|W_2)p(W_2)}{p(w|W_2)p(W_2)+p(w|W_1)p(W_1)}
$$
Using our assumption that the cues are uncorrelated, we have
$$p(w | W) = p(a,v| W) = p(a| A)p(v| V)$$
Under this assumption, the posterior probability reduces to:\footnote{See the appendix for the details of the derivation}
\begin{equation}
 p(W_2 | w)=\frac{1}{1+(1+\epsilon)\exp(\beta_0+\beta_aa+\beta_vv)}
\end{equation}
where 
$$1+\epsilon=\frac{p(W_1)}{p(W_2)}$$
$$\beta_0=\frac{\mu^2_{A2}-\mu^2_{A1}}{2\sigma^2_{A}}+\frac{\mu^2_{V2}-\mu^2_{V1}}{2\sigma^2_{V}}$$

@@

The parameter $\epsilon$ represents the differential between the categories' prior probabilities. However, since the identity of word categories is randomized across participants, $\epsilon$ measures, rather, a response bias to "same" if $\epsilon > 0$, and a response bias to "different" if $\epsilon < 0$. We expect a general bias towards answering "different" because of the categorical nature of our same-different task: When two items are ambiguous but perceptually different, participants might have a slight preference for "different" over "same".

<!--
We fix the values of the means as follows. Since the endpoints of the auditory and the visual continua are the most typical tokens, we assume that the categories are peaked around these endpoints, so we fix the values of the mean to be these endponts.  to be the endpoints of the corresponding continuum, meaning that <!--For example, if both the auditory and visual continua are made of 5 steps going from 0 to 4, then $\mu_{A1}=0$ and $\mu_{A2}=4$ (and similarly $\mu_{V1}=0$, and $\mu_{V2}=4$). --> The values of the means are fixed: They correspond to the value of the most typical tokens in our stimuli. Thus, observations from each modality ($a$ and $v$) are weighted according to their reliability in the posterior: $$\beta_a \propto \frac{1}{\sigma^2_{A}}$$ $$\beta_v \propto \frac{1}{\sigma^2_{V}}$$.

Note that the weights account for categorical variability only. For instance, if the speaker generates a target production $a_t$ from an auditory category
$p(a_t | A) \sim N(\mu_{A}, \sigma^2_{A})$, the ideal model assumes that it has direct access to this production token (i.e., $a=a_t$), and that all uncertainty is about the category membership of this token. However, we might also want to account for noise in the brain and/or in the environment.  For example, the observer might not have access to the exact produced target, but only to the target perturbed by noise. If we assume this noise to be normally distributed, that is,  $p(a | a_t) \sim N(a_t, \sigma^2_{N})$, then integrating over $a_t$ lead to the following simple expression:
$$ p(a | A) \sim N(\mu_{A}, \sigma^2_{A}+\sigma^2_{N})$$
Using an identical reasoning for the case of the visual modality, we end up with the following multimodal weighting scheme in the optimal combination model:

 $$\beta_a \propto \frac{1}{\sigma^2_{A}+\sigma^2_{N_A}}$$ $$\beta_v \propto \frac{1}{\sigma^2_{V} +\sigma^2_{N_V}}$$
To summarize, the posterior provides the optimal model's predictions for how probabilities that characterize uncertainty in each modality can be combined to make categorical decision about the bimodal input. The model will be constructed based on data from the unimodal trials, but the predictions of the optimal model will be compared to human responses in the bimodal trials.

## Auditory and Visual models

Besides the optimal model, we also test the predictions of two baselines models: a visual model, which assumes that participants rely only on visual information, and an auditory model, which assumes that participants rely only on auditory information. If participants rely on both the auditory and the visual modalities, the optimal model would explain more variance in human responses than the visual model alone or the auditory model alone.

## Descriptive model

The visual, auditory, and optimal models are *normative* models. Their predictions are made about human data in the bimodal condition, but their crucial parameters (i.e., variances associated with the visual and auditory modalities) are derived from data in the unimodal conditions.
In addition to these normative models, we consider a *descriptive* model. The parameters of this model are fit to human data in the bimodal condition. If the referential task induces sub-optimality (due, for instance, to the arbitrary nature of the sound-object association), then we predict that the descriptive model should explain more variance than the optimal model does.

>>>>

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

@@

```{r load_packages, include = FALSE}
library(papaja)
library(readr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(knitr)
library(boot)
library(dplyr)
library(langcog)
library(ggthemes)
#theme_set(theme_few())
library(broom)
#library("knitcitations")
#cleanbib()
#options("citation_format" = "pandoc")
```

@@

How does the listener deal with uncertainty to identify the speaker's intended word? One rigorous way to approach this question is through conducting an *ideal observer* analysis. This research strategy provides a characterization of the task/goal and shows what the optimal performance should be under this characterization.\footnote{This strategy is, thus, a general instance of the rational approach to cognition [@anderson90], as well as Marr's computational level of analysis.} When there is uncertainty in the input, the ideal observer performs an optimal probabilistic inference. For example, in order to recognize an ambiguous linguistic input, the model uses all available probabilistic knowledge in order to maximize the accuracy of this recognition. Thus, when the task is well specified, the ideal observer model can be seen as a theoretical upper limit on performance. It is often used, not so much as a realistic model of human performance, as much as a baseline against which human performance can be compared [@Geisler2003]. When there is a deviation from the ideal, this can reveal extra constraints on human cognition, such as limitations on working memory or attentional resources (see Rahnev and Denison, 2018 for further discussion).

The ideal observer analysis has had a tremendous impact not only on speech related research [@Norris08; @clayard08; @feldman2009; @kleinschmidt2015], but also on many other disciplines in the cognitive sciences [for reviews, see @chater06; @Knill04; @tenenbaum11]. In particular, using this research strategy, @clayard08 simulated auditory uncertainty by manipulating the probability distribution of a cue (VOT) that differentiated similar words (e.g., "beach" and "peach"). They found that humans were sensitive to these probabilistic cues and their judgments closely reflected the optimal predictions. In another work, @feldman2009 studied the perceptual magnet effect, which is a phenomenon that involves reduced discriminability near prototypical sounds in the native language (Kuhl, 1991). They showed that this effect can be explained as the consequence of optimally solving the problem of perception under uncertainty. Both @clayard08 and @feldman2009 explored optimal performance under uncertainty in the auditory modality. There is, however, extensive evidence that information from the visual modality, such as the speaker's facial features,  also influences speech understanding [see @Campbell2008 for a review]. @bejjanki2011 offered a mathematical characterization of how cues from speech and lip movements can be optimally combined and showed that human performance during audio-visual phonemic labeling was consistent (at least at the qualitative level) with the behavior of the ideal observer.

This previous reseach, however, did not systematically study speech inderstanding when the visual information is obtained, not through the speaker's facial features, but through the referential context. In fact, experimental findings showed that information about the identity of the semantic referent can be integrated with linguistic information to resolve lexical and syntactic ambiguities in speech [e.g., @Eberhard1995; @Tanenhaus1995; @spivey2002]. To our knowledge, however, no study offered an ideal observer analysis of word identification in such context, that is, when the listener has to combine cues from the sound and the referent. Imagine, for example, that someone is uncertain whether they heard "pea" or "bee", does this uncertainty make them rely more on the referent (e.g., the object being pointed at)? Vice versa, if they are not sure if they saw a bee or a fly, does it make them rely more on the sound? More importantly, when input in both modalities is uncertain to varying degrees, do they weight each modality according to its relative reliability (which is the optimal strategy), or do they over-rely on a particular modality (which is a sub-optimal strategy)?

On the face of it, the question of combining information from the sound and the visual referent might seem similar to that of audio-visual speech integration. Nevertheless, there are at least two fundamental differences between these two cases, and both can influence the way the auditory and visual cues are integrated.
First, in the case of audio-visual speech, modalities may difffer only in terms of their informational reliability. In a referential context, however, modalities also differ in terms of their roles in the referential process: the auditory input represents the *symbol* whereas the visual input represents the *meaning*. It has been suggested that because of its referential property, speech is a privileged signal for humans, starting in infancy [see @vouloumanos2014 for a review].\footnote{There is a debate as to whether speech is privileged for infant and adults for the same reasons. Whereas some researchers suggest that speech is privileged for both infants and adults because of its ability to refer (Waxman et al. XX), others suggest that speech might not have a referential status from the start. Rather, speech might be prefered by infants only because of a low level auditory "overshadowing" (Sloutsky et al. XX).} Thus, in a referential context, it is possible that listeners do not treat the auditory and visual modalities as equivalent sources of information. Instead, there could be a sub-optimal bias for the auditory modality beyond what is expected from informational reliability alone.

<!--An example of such integration is when we determine the size of an object using the visual and haptic modalities [@ernst02], or when we determine the spacial location of a stimulus using the visual and the auditory modalities [@alais04] -->
Second, in the case of audio-visual speech, the auditory and visual stimuli are expected to be perceptually correlated. The expectation for this correlation is such that when there is a mismatch between the auditory and visual input, people still integrate them into a unified (but illusory) percept [@mcgurk1976]. In the case of referential language, however, the multimodal association is by nature *arbitrary* [@saussure1916; @greenberg1957].  For instance, there is no logical/perceptual connection between the sound "bee" and the corresponding insect. Moreover, variation in the way the sound "bee" is pronounced is generally not expected to correlate perceptually with variation in the shape (or any other visual property) in the category of bees. In sum, cue combination in the case of arbitrary audio-visual associations (word-referent) is likely to be less automatic, more effortful, and therefore less conducive to optimal integration than it is in the case of perceptually correlated associations (as in audio-visual speech perception).

In the current study, we investigate how people combine cues from the auditory and the visual modality to recognize words in a referential context. In particular, we study how this combination is performed under various degrees of uncertainty in both the auditory and the visual modality.  We perform a rational analysis of the task. First we propose an ideal observer model that performs the combination in an optimal fashion. Second we compare the predictions of the optimal model to human responses. Humans can deviate from the ideal for several reasons. For instance, as mentioned above, a sub-optimality can be induced by the suggested privileged status of speech or by the arbitrariness of the referential association.  In order to study possible patterns of sub-optimality, we compare the optimal model (which provides a normative benchmark) to a descriptive model (which is fit to human responses). Comparing parameter estimates between these two formulations allows us to quantify the degree of deviation from optimality.

<!--We do not perform the ideal observer analysis at the individual level, but at the population level. There might be some individual variation, but this study addresses the overall performance in the population as a whole.-->

@@

# Pardigm and Models

In this section we, first, briefly introduce the multimodal combination task. Then we explain how behavior in this paradigm can be characterized optimally with an ideal observer model.

@@

We introduce a new task that tests word recognition in a referential context.  We use two visual categories (cat and dog) and two auditory categories (/b/ and /d/ embedded in the minimal pair /aba/-/ada/). For each participant, an arbitrary pairing is set between the auditory and the visual categories, leading to two audio-visual word categories (e.g., dog-/aba/, cat-/ada/). In each trial, participants are presented with an audio-visual target (the prototype of the target category), immediately followed by an audio-visual test stimulus (Figure\ \@ref(fig:task)). The test stimulus may differ from the target in both the auditory and the visual components.  After these two presentations, participants press "same" or "different."

@@

This paradim is adapted from a task by @sloutsky2003, in which participants are asked whether or not the two audio-visual presentations are *identical*. The original task has been used with both chidlren and adults to probe  audio-visual encoding [see @robinson2010 for a review]. We are rather interested in categorization responses, i.e., determining whether or not two similar tokens are members of the same phonological/semantic category. Therefore, testing in our task is category-based: Participants are asked to press "same" if they think the second item (the test) belonged to the same category as the first (target) (e.g.,  dog-/aba/), even if there is a slight difference in the sound, in the referent, or in both. They are instructed to press "different" only if they think that the second stimulus was an instance of the other word category (cat-/ada/). The task also includes trials where pictures are hidden (audio-only) or where sounds are muted (visual-only). These unimodal trials provide us with the participants' evaluation of the probabilistic information present in the auditory and visual categories. This unimodal distributions are used as inputs to the optimal combination model, described below.

@@

We construct an ideal observer model that combines probabilistic information from the auditory and visual modalities. In contrast to the  model used in most research on multisensory integation [@ernst02; @alais04]---which typically studies continuous stimuli (e.g., size, location)---this probabilistic information cannot be characterized with sensory noise/variability, only.  Indeed, our task involves repsonses over categorical variables (phonemes and concepts), and therefore, the optimal model in our case should take into account, not only the noise variability around an individual perceptual estimate, but also its *within category variability*, i.e., the uncertainty related to whether this perceptual estimate belongs to a given category [see also @Bankieris17; @bejjanki2011]. In what follows, we describe a model that accounts for both type of varibility. First, we describe the model in the simplified case of categorical variability only. Second, we augment this simplified model to account for sensory noise.

### Categorical variability
We assume that both the auditory categories (i.e., /aba/ and /ada/) and the visual categories (cat and dog) are distributed along a single acoustic and semantic dimension, respectively (Figure\ \@ref(fig:model)). Moreover, we assume that all categories are normally distributed. Formally speaking, if $A$ denotes an auditory category (/ada/ or /aba/), then the probability that a point $a$ along the acoustic dimension belongs to the category $A$ is
$$ p(a | A) \sim  N(\mu_A, \sigma^2_A) $$
where $\mu_A$ and $\sigma^2_A$ are respectively the mean and the variance of the auditory category.
Similarly, the probability that a point $v$ along the visual dimension belongs to the category $V$ is
$$ p(v | V) \sim  N(\mu_V, \sigma^2_V) $$
where $\mu_V$ and $\sigma^2_V$ are the mean and the variance of the visual category.
An audio-visual signal $w=(a,v)$ can be represented as a point in the audio-visual space. These tokens belong to bivariate distributions in this bi-dimentional sapce. We call these bivariate distributions *Word categories*, noted $W$, and are distributed as follows: 
$$ p(w | W) \sim  N(M_W, \Sigma_W) $$
where $M_W=(\mu_A, \mu_V)$ and $\Sigma_W$ are the mean and the covariance matrix of the word category. The main assumption of the model is that the auditory and visual variables are independent (i.e., uncorrelated), so the covariance matrix is simply:
 \[
   \Sigma_W=
  \left[ {\begin{array}{cc}
   \sigma^2_A & 0 \\
   0 & \sigma^2_V \\
  \end{array} } \right]
\]

```{r model, out.width = "\\textwidth", fig.pos = "!h", fig.cap = "Illustration of the model using simulated data. A word category is defined as the joint bivariate distribution of an auditory category (horizontal, bottom panel) and a visual semantic category (vertical, left panel). Upon the presentation of a word token w, participants guess whether it is sampled from the word type W1 or from the word type W2. Decision threshold is where the guessing probability is 0.5."}
knitr::include_graphics("pictures/model.png", dpi = 108)
```

@@

Now we turn to the crucial question of modeling how optimal decision should proceed given the probabilistic (categorical) information in the auditory and the visual modalities, as characterized above. We have two word categories: dog-/aba/ ($W_1$) and cat-/ada/ ($W_2$).\footnote{This mapping is randomized in the experiments.} When making decisons, participants can be understood as choosing one of these two word categories (Figure\ \@ref(fig:model)). For an ideal observer, the probability of choosing category 2 when presented with an audio-visual instance $w=(a,v)$ is the posterior probability of this category:
$$
p(W_2 | w)=\frac{p(w|W_2)p(W_2)}{p(w|W_2)p(W_2)+p(w|W_1)p(W_1)}
$$
Using our assumption that the cues are uncorrelated, we have
$$p(w | W) = p(a,v| W) = p(a| A)p(v| V)$$
Under this assumption, the posterior probability reduces to:\footnote{See the appendix for the details of the derivation}
\begin{equation}
 p(W_2 | w)=\frac{1}{1+(1+b)\exp(\beta_0+\beta_aa+\beta_vv)}
\end{equation}
where 
$$1+b=\frac{p(W_1)}{p(W_2)}$$
$$\beta_0=\frac{\mu^2_{A2}-\mu^2_{A1}}{2\sigma^2_{A}}+\frac{\mu^2_{V2}-\mu^2_{V1}}{2\sigma^2_{V}}$$

@@

The parameter $b$ represents the differential between the categories' prior probabilities. However, since the identity of word categories is randomized across participants, $b$ measures, rather, a response bias to "same" if $b > 0$, and a response bias to "different" if $b < 0$. We expect a general bias towards answering "different" because of the categorical nature of our same-different task: When two items are ambiguous but perceptually different, participants might have a slight preference for "different" over "same". As for the means, their values are fixed, and they correspond to the most typical tokens in our stimuli. Thus, observations from each modality ($a$ and $v$) are weighted according to their reliability in the posterior: $$\beta_a \propto \frac{1}{\sigma^2_{A}}$$ $$\beta_v \propto \frac{1}{\sigma^2_{V}}$$.

### Sensory variability

So far, we only accounted for categorical variability. For instance, if the speaker generates a target production $a_t$ from an auditory category
$p(a_t | A) \sim N(\mu_{A}, \sigma^2_{A})$, the ideal model assumes that it has direct access to this production token (i.e., $a=a_t$), and that all uncertainty is about the category membership of this token. However, we might also want to account for noise in the brain and/or in the environment.  For example, the observer might not have access to the exact produced target, but only to the target perturbed by noise. If we assume this noise to be normally distributed, that is,  $p(a | a_t) \sim N(a_t, \sigma^2_{N_A})$, then integrating over $a_t$ leads to the following simple expression:
$$ p(a | A) \sim N(\mu_{A}, \sigma^2_{A}+\sigma^2_{N_A})$$
Similarly, in the case of sensory noise in the visual modality, we get
$$ p(a | V) \sim N(\mu_{V}, \sigma^2_{V}+\sigma^2_{N_V})$$
Finally, using exactly the same derivation as above, we end up with the following multimodal weighting scheme in the optimal combination model (Equation 1) which takes into account both categorical and sensory variability:

 $$\beta_a \propto \frac{1}{\sigma^2_{A}+\sigma^2_{N_A}}$$ $$\beta_v \propto \frac{1}{\sigma^2_{V} +\sigma^2_{N_V}}$$

### Optimal cue combination

The posterior (Equation 1) provides the optimal model's predictions for how probabilities that characterize uncertainty in the auditory and the viusal modalities can be combined to make categorical decisions. Estimates of the probability distributions in each modality are derived by fitting logistic regressions to participants' responses in the unimodal conditions, where participants only hear the sounds or only see the pictures (Figure\ \@ref(fig:task)) \footnote{Further technical detail about model fitting in the unimodal conditions will be given in the method section of Experiment 1}. Using these derived parameters, the optimal model makes predictions about responses in the bimodal condition where participants both hear the sounds and see the pictures.

### Auditory and Visual baselines

The predictions of the optimal model will be compared to two baselines. The first basline is a visual model which assumes that participants rely only on visual information, and the auditory model, which assumes that participants rely only on auditory information. More precisely, these baseline models assumes that the participants' respones in the bimodal condition will not be different from their response in one of the unimodal conditions. If, as we expect, the participants rely on both the auditory and the visual modalities to make decision in the bimodal condition, the optimal model would explain more variance in human responses than the visual model alone or the auditory model alone.

## Descriptive model and analysis of sub-obtimality

The optimal models (as well as the auditory and visual baselines) are *normative* models. Their predictions are made about human data in the bimodal condition, but their crucial parameters (i.e., variances associated with the visual and auditory modalities) are derived from data in the unimodal conditions.
In addition to these normative models, we consider a *descriptive* model. The parameters of this model are fit to actual respnses in the bimodal condition. If the referential task induces sub-optimality (due, for instance, to the arbitrary nature of the sound-object association), then the descriptive model should explain more variance than the optimal model does.
